I Compensate, therefore I Am \\ (accurate for DNN verification)

### 2024-01-19 Paper/introduction.tex
introduction

In this paper, we focus on DNNs trained in a "natural", 
%uncovering what makes the DNNs trained in a natural way so hard to verify (
because for "easier" DNNs, adequate methods already exist. 
To do so, we analyse the abstraction mechanisms at the heart of several efficient algorithms, namely Eran-DeepPoly, the Linear Programming approximation, PRIMA, and different versions of ($\alpha$)($\beta$)-CROWN. All these algorithms compute lower or/and upper bounds for the values of neurons (abstraction on values) for inputs in the considered input region, and conclude based on such bounds. For instance, if for all image $I'$ in the neighbourhood of image $I$, we have $weight_{I'}(n'-n) < 0$ for $n$ the output neuron corresponding to the expected class, then we know that the DNN is robust in the neighbourhood of image $I$. We restrict the formal study to DNNs using only the standard ReLU activation function, although nothing specific prevents the results to be extended to more general architectures. We uncover that {\em compensations} (see next paragraph) is the phenomenon creating inaccuracies. We verified experimentally that indeed DNNs trained in a natural way have much more heavy compensating pairs than DNNs trained in a robust way.

@@

This paper does not focus on producing the most efficient tool, and we did not spend engineering efforts to optimize it. The focus is instead on the novel notion of compensation, the associated methodology and its evaluation. For instance, our implementation is fully in Python, with uncompetitive runtime for our DeepPoly implementation ($\approx 100$ slower than in $\beta$-CROWN). Still, evaluation of the methodology versus even the most efficient tools reveals a lot of potential for the notion of compensation, opening up several opportunities for applying it in different contexts of DNN verification (see Section \ref{Discussion}).

@@

\noindent {\bf Comparison with related work:} Here, we will compare with some (but not all) main verification tools for DNNs, to better explain our methodology and how it differs with the existing SOTA.
Compared with the exact encoding of a DNN using MILP \cite{MILP}, our algorithm can be seen as a way to help the MILP tool by telling it to not spend time to branch on some ReLU nodes which have low impact on the particular node we are treating, at the cost of a small inaccuracy. If the ReLU nodes to abstract are chosen accurately, the result should be more accurate than early stopping the MILP algorithm using all the nodes.

Compared with the linear relaxation of the MILP encoding, our algorithm is strictly more accurate by design, but it will also be slower.

@@

Concerning PRIMA, the idea is to keep explicitly dependencies between neurons, computing bounds layer by layer (as we do). This allows to keep very efficiently dependencies from potentially many layers beforehand. We take care of dependencies between neurons in a different way, as compensation is the reason why there are dependencies between neurons. 
Our method is more accurate locally, but we will tend to lose precision for dependencies created many layers ago. Experimental results tend to show that most of the dependencies are local, in the few last layers (because ReLU nodes will likely clip those that happened many layers ago). Also, the dependencies between nodes limit the parallelism, unlike in our method, which explains why we obtain both faster and more accurate results than PRIMA.

@@

Finally, algorithms abstracting the network (e.g. Reluplex / Marabou \cite{Reluplex,katz2019marabou}) are very different from algorithms abstracting the values (PRIMA, ($\alpha$)($\beta$)-CROWN)\cite{prima,crown} $\cdots$ These algorithms have been developed to be complete, so they are much slower but also more accurate than what we propose.

>>>>

In this paper, we focus on DNNs trained in a "natural", 
%uncovering what makes the DNNs trained in a natural way so hard to verify (
because for "easier" DNNs, adequate methods already exist. 
To do so, we analyse the abstraction mechanisms at the heart of several efficient algorithms, namely Eran-DeepPoly \cite{deeppoly}, the Linear Programming approximation \cite{MILP}, PRIMA \cite{prima}, and different versions of ($\alpha$)($\beta$)-CROWN \cite{crown}. All these algorithms compute lower or/and upper bounds for the values of neurons (abstraction on values) for inputs in the considered input region, and conclude based on such bounds. For instance, if for all image $I'$ in the neighbourhood of image $I$, we have $weight_{I'}(n'-n) < 0$ for $n$ the output neuron corresponding to the expected class, then we know that the DNN is robust in the neighbourhood of image $I$. We restrict the formal study to DNNs using only the standard ReLU activation function, although nothing specific prevents the results to be extended to more general architectures. We uncover that {\em compensations} (see next paragraph) is the phenomenon creating inaccuracies. We verified experimentally that indeed DNNs trained in a natural way have much more heavy compensating pairs than DNNs trained in a robust way.

@@

This paper does not focus on producing the most efficient tool, and we did not spend engineering efforts to optimize it. The focus is instead on the novel notion of compensation, the associated methodology and its evaluation. For instance, our implementation is fully in Python, with uncompetitive runtime for our DeepPoly implementation ($\approx 100$ slower than in CROWN). Still, evaluation of the methodology versus even the most efficient tools reveals a lot of potential for the notion of compensation, opening up several opportunities for applying it in different contexts of DNN verification (see Section \ref{Discussion}).

@@

\noindent {\bf Comparison with related work:} Here, we will compare with several (but not all) main verification tools for DNNs, to better explain our methodology and how it differs with the existing SOTA. Compared with the exact encoding of a DNN using MILP \cite{MILP}, our algorithm can be seen as a way to help the MILP tool by telling it to not spend time to branch on ReLU nodes which have low impact on the particular node we are treating, at the cost of a small inaccuracy. If the ReLU nodes to abstract are chosen accurately, the result should be more accurate than early stopping the MILP algorithm using all the nodes.

Compared with the linear relaxation of the MILP encoding \cite{MILP}, our algorithm is strictly more accurate by design, but it will also be slower.

@@

Concerning PRIMA \cite{prima}, the idea is to keep explicitly dependencies between neurons, computing bounds layer by layer (as we do). This allows to keep very efficiently dependencies from potentially many layers beforehand. We take care of dependencies between neurons in a different way, as compensation is the reason why there are dependencies between neurons. 
Our method is more accurate locally, but we will tend to lose precision for dependencies created many layers ago. Experimental results tend to show that most of the dependencies are local, in the few last layers (because ReLU nodes will likely clip those that happened many layers ago). Also, the dependencies between nodes limit the parallelism, unlike in our method, which explains why we obtain both faster and more accurate results than PRIMA.

@@

Finally, algorithms abstracting the network (e.g. Reluplex / Marabou \cite{Reluplex,katz2019marabou}) are very different from algorithms abstracting the values (PRIMA, ($\alpha$)($\beta$)-CROWN)\cite{prima,crown}$\ldots$ These algorithms have been developed to be complete, so they are much slower but also more accurate than what we propose.
